### 2016-06-18 ch02-qlogms/excitation-design.tex
Some diction suggestions
Some recommendations private defense
Add some TODOs from Jury/self

\section{Introduction}
\label{sec:excitation:intro}
Input design deals with the task of choosing an excitation signal in a way so that the information obtained from the experiment is maximized.  
This is a problem which has a long history in statistics. 
In system identification, the focus has mostly been on input design for dynamic systems.
A common approach is to design the input spectrum and thereby maximizing the information matrix of the estimates, see e.g. \citep{Fedorov1972,Goodwin1977}.

@@

First, we look at the band-limited $1/f$ power spectrum given by
\begin{align*}
\label{eq:excitation:pink}
\phi_u(\omega) &=\begin{cases}\frac{1}{|\omega|} &\text{if $|\omega|\in[\wn \kMin\; \wn \kMax]$,}\\
                        0           &\text{otherwise}.
            \end{cases}
\end{align*}
With this input spectrum, the power inside $\wdB \approx \ClosedInterval{(1-\damping)\wn, (1+\damping)\wn}$ is
\begin{align}
  \frac{1}{2\pi}
    \int\limits_{(1-\damping)\wn}^{(1+\damping)\wn}
     \frac{1}{\omega} \dd{\omega}
  = \frac{1}{2\pi}
    \ln{\frac{1+\damping}{1-\damping}},
\end{align}
which is independent of $\wn$. The robustness of the $1/f$ input has been noted previously by e.g. \citet{Rojas2007,Rojas2012,Goodwin2006GBO}.

@@

\subsection{Sums of Second Order Systems}
This section extends the previous section to more general systems than second order.
Consider a system
\begin{align}
G(s,\theta) &= \sum_{i=1}^{K} G_i(s,\theta_i)
             =\sum_{i=1}^{K}  \frac{1}
                                   {   \frac{s^2}{\wn[i]^2}
                                    + 2\frac{\damping[i]}{\wn[i]}s
                                    + 1},\\
\theta_i&=[\wn[i]\;\damping[i]]^{\T}.
\label{eq:excitation:sumSys}
\end{align}
The systems  $G_i$ will be called the \emph{sub-systems} of $G$. 
Without loss of generality, we assume that $\wn[1] < \wn[2] < \cdots < \wn[K]$. 
To properly identify the system $G$, excitation is required such that all subsystems are properly excited, regardless of their individual resonance frequencies.

@@

In \figref{fig:excitation:coupling} the largest element in $\partial G_{1,2}$ is plotted for different dampings and separations. It is clear that highly resonant systems have a large degree of decoupling and that resonances that are well-separated increases the decoupling. 
The diagonal blocks of the normalized $\Fi{\theta}$ in all cases have some elements close to $1$. 
In this example, the $\damping$ is set to be the same in both sub-systems. 
Using different dampings in the sub-systems would, however, not change the argument, the systems with the lowest damping will limit the separation.

@@

  Therefore a signal with an equidistant frequency grid is undesirable for
  our situation as it wastes signal power at high frequencies at the cost of an increased variance in the lower side of the band.

@@

  \begin{property}\label{prop:excitation:log:relBW}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
    Note that this is independent of the frequency $f$.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} and programming languages, it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$) instead of the grid ratio $\alpha$.
    Obviously, such alternative specifications are equivalent.
  \end{remark}
  \begin{example} \label{eg:excitation:logarithmicGrid:matlab}
    The \MATLAB code \mcode[mathescape]{f = logspace(log10($f_{\min}$), log10($f_{\max}$), $F$)} produces a logarithmically spaced frequency vector \mcode{f} with $F$ elements in the range $\left[ f_{\min}, f_{\max} \right]$.
    As such, the effective frequency ratio $\alpha = \sqrt[F]{\frac{f_{\max}}{f_{\min}}}$ in this vector.
  \end{example}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

@@

  Finally, $\C_{X}$ denotes the covariance matrix of $X\left(\omega\right) = G\left(\omega, \theta_0 \right) - \frac{Y_0\left(\omega \right) + E\left( \omega \right)  }{ U\left( \omega \right) } $, evaluated at the excited frequencies $\omega_{\excited}$.
  By assuming that the $X\left( \omega \right) $ are independently distributed over the frequency $\omega$, the variance-covariance matrix $\C_{X}$ is a diagonal matrix \citep{Pintelon2012}:
  \begin{equation}
    \C_{X} =
               \left[ \left.
               \begin{matrix}
                 \frac{\sigma^2_E\left(\omega_1 \right) }
                      {\abs{ U\left(\omega_1 \right)}^2 } & \deemph{0} & \deemph{0}\\
                 \deemph{0} & \ddots & \deemph{0} \\
                 \deemph{0} & \deemph{0} & \frac{\sigma^2_E\left(\omega_F \right) }
                              {\abs{U\left(\omega_F \right)}^2 }
               \end{matrix}
               \right.^{\vphantom{x}}\right]
    \text{.}
  \end{equation}

@@

  Now we can interpret \figref{fig:excitation:worstCaseBeta} in more detail.
  We see that for $\beta \in \set{0,1}$ (i.e. perfect alignment), the variance of a course grid (full lines) is lower than it is for a fine grid (dashed lines).
  This is to be expected as a coarse grid places more power at each individual line.
  Thereby it yields a smaller variance at those lines.
  It is better tailored to systems which are closely aligned to the excited frequency grid.
  However, there is a critical consequence: when there is a moderate misalignment $\beta$ of the system and the excitation grid, the variance increases significantly.
  This effect is more important for lightly-damped systems, as their $3\unit{dB}$ bandwidth is very narrow and hence their main response is easier to `miss'.

>>>>

%TODO: define robustness upon first use
\section{Introduction}
\label{sec:excitation:intro}
Input design deals with the task of choosing an excitation signal in a way so that the information obtained from the experiment is maximized.  
This is a problem which has a long history in statistics. 
In system identification, the focus has mostly been on input design for dynamic systems.
A common approach is to design the input spectrum and thereby maximizing the information matrix of the estimates, see e.g. \citep{Fedorov1972,Goodwin1977}.

@@

First, we look at the band-limited $1/f$ power spectrum given by
\begin{align*}
\label{eq:excitation:pink}
\phi_u(\omega) &=\begin{cases}\frac{1}{|\omega|} &\text{if $|\omega|\in[\wn \kMin\; \wn \kMax]$,}\\
                        0           &\text{otherwise}.
            \end{cases}
\end{align*}
With this input spectrum, the power inside $\wdB \approx \ClosedInterval{(1-\damping)\wn, (1+\damping)\wn}$ is
\begin{align}
  \frac{1}{2\pi}
    \int\limits_{(1-\damping)\wn}^{(1+\damping)\wn}
     \frac{1}{\omega} \dd{\omega}
  = \frac{1}{2\pi}
    \ln{\frac{1+\damping}{1-\damping}},
\end{align}
which is independent of $\wn$. 
The robustness of the $1/f$ input has been noted previously by e.g. \citet{Rojas2007,Rojas2012,Goodwin2006GBO}.
%TODO: please define the meaning of "robustness"

@@

\subsection{Sums of Second Order Systems}
This section extends the previous section to more general systems than second order.
Consider a system
\begin{align}
G(s,\theta) &= \sum_{i=1}^{K} G_i(s,\theta_i)
             =\sum_{i=1}^{K}  \frac{1}
                                   {   \frac{s^2}{\wn[i]^2}
                                    + 2\frac{\damping[i]}{\wn[i]}s
                                    + 1},\\
\theta_i&=[\wn[i]\;\damping[i]]^{\T}.
\label{eq:excitation:sumSys}
\end{align}
The systems  $G_i$ will be called the \emph{sub-systems} of $G$. 
Without loss of generality, we assume that $\wn[1] < \wn[2] < \cdots < \wn[K]$. 
To properly identify the system $G$, excitation is required such that all subsystems are properly excited, regardless of their individual resonance frequencies.
%TODO: very limited model (K=1). Comment.

@@

In \figref{fig:excitation:coupling} the largest element in $\partial G_{1,2}$ is plotted for different dampings and separations. It is clear that highly resonant systems have a large degree of decoupling and that resonances that are well-separated increases the decoupling. 
The diagonal blocks of the normalized $\Fi{\theta}$ in all cases have some elements close to $1$. 
In this example, the $\damping$ is set to be the same in both sub-systems. 
%TODO: !! (EG) Comment on that this is not a requirement
Using different dampings in the sub-systems would, however, not change the argument, the systems with the lowest damping will limit the separation.

@@

  Therefore a signal with an equidistant frequency grid is undesirable for our situation as it wastes signal power at high frequencies at the cost of an increased variance in the lower side of the band.
  %TODO: Not clear at this stage

@@

  \begin{property}\label{prop:excitation:log:relBW}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
    Note that this is independent of the frequency $f$.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} and programming languages, it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$) instead of the grid ratio $\alpha$.
    Obviously, such alternate specifications are equivalent.
  \end{remark}
  \begin{example} \label{eg:excitation:logarithmicGrid:matlab}
    The \MATLAB code \mcode[mathescape]{f = logspace(log10($f_{\min}$), log10($f_{\max}$), $F$)} produces a logarithmically spaced frequency vector \mcode{f} with $F$ elements in the range $\left[ f_{\min}, f_{\max} \right]$.
    As such, the effective frequency ratio $\alpha = \sqrt[F]{\frac{f_{\max}}{f_{\min}}}$ in this vector.
  \end{example}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

@@

  Finally, $\C_{X}$ denotes the covariance matrix of $X\left(\omega\right) = G\left(\omega, \theta_0 \right) - \frac{Y_0\left(\omega \right) + E\left( \omega \right)  }{ U\left( \omega \right) } $, evaluated at the excited frequencies $\omega_{\excited}$.
  By assuming that the $X\left( \omega \right) $ are independently distributed over the frequency $\omega$, the variance-covariance matrix $\C_{X}$ is a diagonal matrix \citep{Pintelon2012}:
  \begin{equation}
    \C_{X} =
               \left[ \left.
               \begin{matrix}
                 \frac{\sigma^2_E\left(\omega_1 \right) }
                      {\abs{ U\left(\omega_1 \right)}^2 } & \deemph{0} & \deemph{0}\\
                 \deemph{0} & \ddots & \deemph{0} \\
                 \deemph{0} & \deemph{0} & \frac{\sigma^2_E\left(\omega_F \right) }
                              {\abs{U\left(\omega_F \right)}^2 }
               \end{matrix}
               \right.^{\vphantom{x}}\right]
    \text{.}
  \end{equation}
  \begin{remark}
  To introduce only limited prior knowledge into the input design, we shall assume that the noise is white.
  Hence, $\sigma^2_E(\omega) = \sigma^2_E$ is constant over the whole frequency band of interest.
  If the noise coloring is known beforehand (or measured in a previous experiment), the design can be adapted in a straightforward manner.
  \end{remark}

@@

  Now we can interpret \figref{fig:excitation:worstCaseBeta} in more detail.
  We see that for $\beta \in \set{0,1}$ (i.e. perfect alignment), the variance of a course grid (full lines) is lower than it is for a fine grid (dashed lines).
  This is to be expected as a coarse grid places more power at each individual line.
  Thereby it yields a smaller variance at those lines.
  It is better tailored to systems that are closely aligned to the excited frequency grid.
  However, there is a critical consequence: when there is a moderate misalignment $\beta$ of the system and the excitation grid, the variance increases significantly.
  This effect is more important for lightly-damped systems, as their $3\unit{dB}$ bandwidth is very narrow and hence their main response is easier to `miss'.
